---
title: "Midterm 1-Note Sheet"
author: "Jaskirat"
date: "2024-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#LAB 1
Getting working directory: 
```{r}
# getwd()
```
Making a vector of values:
```{r}
#x <- c(4,6,8,7,7,7,8)
```
Talks about R markdown

#LAB 2
Getting the sum of variables 
```{r}
treatment <- 36
control <- 10
sum <- sum(treatment, control)
```
Finding the class of a variable
```{r}
# class(variable name)
# is.integar(variable name) ....to see if varibale name is a interger or not
# as.integar(variable name) .... creat a varibale specified to be an integar

```
Finding NAs in a data set
```{r}
# is.na(variable) or can use anNA(variable)
# use na.rm=T to removes NAs from a data set when doing arthermitic 
```
 Part 2: Creating a vector 
```{r}
day_of week <- c("Monday","Tuesday", "Wednesday")
# my_vector_sequence <- c(1:100)  .....to generate a sequence of numbers 
#day_of_week[2] would pull out the second value in the vector
```
 Creating a matrix
```{r}
#harry_potter_matrix <- matrix(vector name, nrow=8, byrow=T)
#naming columns
    #colnames(harry_potter_matrix) <- region
    #rownames(harry_potter_matrix) <- titles
```
Data matrix analysis
```{r}
#global <- rowSums(harry_potter_matrix)
#create a new row in the data column
       #harry_potter_matrix <- cbind(harry_potter_matrix, global)
#harry_potter_matrix[2,1].....get second row, first column 
#second <- harry_potter_matrix[ ,2]......select values of the entire second row and assign to new variable

#rowMeans() can be used to get the mean of rows. 
```

#LAB 3
Making a data frame from raw data
```{r}
#Sex <- c("male", "female", "male")
#Length <- c(3.2, 3.7, 3.4)
#Weight <- c(2.9, 4.0, 3.1)
#hbirds <- data.frame(Sex, Length, Weight)
```
Getting info of a data frame
```{r}
#names(data_frame).....gives column names
#dim(data_frame).....gives dimensions of data frame
#glimpse(data_frame)
#summary(data_frame)
#nrows(data_frame).....get number of rows 
#ncol(data_frame).....gives number of columns
#head(data_frame)....prints first few rows of a data frame
#tail(data_frame)....prints last few rows of a data frame
#table(data_frame$variable).....gives counts of the number of observations
#str(data_frame).....gives structure of data frame
#hbirds[ ,2]....gives info on the second column 
#hbirds$column_name.......gets all the info from a specific column variable
```
Changing names of columns to lowercase
```{r}
#hbirds <- data.frame(sex=Sex, length=Length, weight_g=Weight)
```
Adding columns to a data frame
```{r}
#new_bird <- c("female", 3.6, 3.9)
#hbirds<- rbind(hbirds, new_bird).....adds data to an existing column 

#hbirds$neighborhood <- c("lakewood", "brentwood", "lakewood", "scenic Heights")....adds a new column 
```
Writing a csv file
```{r}
#write.csv(hbirds, "hbirds_data.csv", row.names = FALSE)
```
Part2: 
importing data
```{r}
#variable_name <- read_csv("filename.csv")

#changing variable to a factor
#dataframe$column_name <- as.factor(dataframe$column_name)
#levels(dataframe$column_name)
```
pulling out observations
```{r}
#filter(data_frame, length<30)
```
Finding sleepiest animal in a dataframe
```{r}
#max(sleep$sleep_total)
#filter(sleep, sleep_total==19.9)
```

#LAB 4
Select function
```{r}
#select(data_frame, "Variable1, "Variable2")....get specific columns
#select(data_frame, variable1:variable5).....to get a range of columns
#select(fish, -"fish_id", -"annnumber", -"length").....select everything expect these names
#select(fish, contains("length")).....select everything with a specific word in it
# ends_with() = Select columns that end with a character string  
# contains() = Select columns that contain a character string  
# matches() = Select columns that match a regular expression  
# one_of() = Select columns names that are from a group of names 
# starts_with()
#select_if(fish, is.numeric).....select if a specific class of variables
#select_if(fish, ~!is.numeric(.))....select if it is not a numeric 


#select_all(mammals, tolower).....change column names to lowercase...use toupper for uppercase
```
Renaming column names 
```{r}
#mammals_new <- rename(mammals, genus="Genus", wean_mass="wean mass", afr="AFR", max_life="max. life", litters_size="litter size", litters_per_year="litters/year")
```
Janitor library
```{r}
#library(janitor)
#clean_names(mammals)....to get clean names
```
Filter Function
```{r}
#filter(fish, length %in% c(167, 175)).....filter multiple values in the same variable
#filter(fish, between(scalelength, 2.5, 2.55)).....for a range of values
#filter(fish, near(radii_length_mm, 2, tol = 0.2))... filter if its close to a specific values
#filter(fish, lakeid == "AL" & length > 350)....AND function
#filter(fish, lakeid == "AL" | length > 350)....OR function
#`filter(condition1, !condition2)` will return all rows where condition one is true but condition 2 is not.
#`filter(xor(condition1, condition2)` will return all rows where only one of the conditions is met, and not when both conditions are met.

#Example:
#filter(fish, length > 400, (scalelength > 11 | radii_length_mm > 8))
```

#LAB 5
Pipes
```{r}
#mammals %>% 
  #select(family, genus, species, gestation) %>% 
  #filter(gestation>=15)
```
Arrange command
```{r}
#arrange(desc(radii_length_mm))
```
Mutate command
```{r}
#fish %>% 
  #mutate(length_mm = length*10) %>% 
  #select(fish_id, length, length_mm)


#change all enteries to lowercase
#mammals %>%
  #mutate_all(tolower)

#change specific columns
#mammals %>% 
  #mutate(across(c("order", "family"), tolower))

#replacing speicifc columns to NA values
#mammals %>% 
  #select(genus, species, newborn) %>%
  #mutate(newborn_new = ifelse(newborn == -999.00, NA, newborn))%>% 
  #arrange(newborn)


```

#LAB 6
Getting proportions
```{r}
#must use janitor package first
#tabyl(superhero_info, Alignment)
#superhero_info %>% 
  #tabyl(Publisher)
```

#LAB 7
```{r}
# library("skimr")......used to produce summaries of data

```
Making histograms of data
```{r}
#hist(msleep24$sleep_total_24)
```
Summarize function:
```{r}
#msleep %>% 
  #filter(bodywt > 200) %>% 
  #summarize(mean_sleep_large= mean(sleep_total))

#msleep %>% 
  #filter(order == "Primates") %>% 
  #summarize(mean_primates = mean(bodywt), min_primates = min(bodywt), max_primates = max(bodywt), total=n())


```
n_distinct function:
```{r}
#msleep %>% 
  #summarize(n_genra=n_distinct(genus)). #this is going to count the number of genera in msleep


#data_frame %>% distinct(name)
#unique(data_frame$name)
```
Finding number of NAs in each group:
```{r}
#penguins %>% 
  #group_by(island) %>% 
  #summarize(number_NAs=sum(is.na(body_mass_g)))

#Removing NAs
#penguins %>% 
  #filter(!is.na(body_mass_g)) %>% #pulls out all the observations with a number 
  #group_by(island) %>% 
  #summarize(mean_body_mass=mean(body_mass_g))


```
count function
```{r}
#penguins %>% 
  #count(island, sort = T)


#penguins %>% 
  #count(island, species, sort = T).....counting across multiple variables
```





