---
title: "Homework 7"
author: "Jaskirat Singh"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
library(naniar)
``` 

## Data
**1. For this homework, we will use two different data sets. Please load `amniota` and `amphibio`.**  

`amniota` data:  
Myhrvold N, Baldridge E, Chan B, Sivam D, Freeman DL, Ernest SKM (2015). “An amniote life-history
database to perform comparative analyses with birds, mammals, and reptiles.” _Ecology_, *96*, 3109.
doi: 10.1890/15-0846.1 (URL: https://doi.org/10.1890/15-0846.1).
```{r}
amniota <- read_csv("data/amniota.csv")
```

`amphibio` data:  
Oliveira BF, São-Pedro VA, Santos-Barrera G, Penone C, Costa GC (2017). “AmphiBIO, a global database
for amphibian ecological traits.” _Scientific Data_, *4*, 170123. doi: 10.1038/sdata.2017.123 (URL:
https://doi.org/10.1038/sdata.2017.123).
```{r}
amphibo <- read_csv("data/amphibio.csv")
```

## Questions  
**2. Do some exploratory analysis of the `amniota` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  
```{r}
glimpse(amniota)
```

**3. Do some exploratory analysis of the `amphibio` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  
```{r}
glimpse(amphibo)
```

**4. How many total NA's are in each data set? Do these values make sense? Are NA's represented by values?**   
```{r}
amniota %>% 
  map_df(~ sum(is.na(.)))
```
```{r}
amphibo %>% 
  map_df(~ sum(is.na(.)))
```
When looking at the number of NA's in the amniota it seems that there are no NA values however based on the glimpse data above, it was seen that they are written as -999 in the data instead. When looking at the amphibo data, there appears to be rocrded as NA values and can be seen from the above code.


**5. Make any necessary replacements in the data such that all NA's appear as "NA".**   
```{r}
amniota2 <- read_csv(file = "data/amniota.csv", 
                  na = c("NA", " ", ".", "-999"))
```

**6. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amniota` data.**  
```{r}
miss_var_summary(amniota2)
```

**7. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amphibio` data.**
```{r}
miss_var_summary(amphibo)
```

**8. For the `amniota` data, calculate the number of NAs in the `egg_mass_g` column sorted by taxonomic class; i.e. how many NA's are present in the `egg_mass_g` column in birds, mammals, and reptiles? Does this results make sense biologically? How do these results affect your interpretation of NA's?**  
```{r}
amniota2 %>%
  group_by(class) %>% 
  select(class, egg_mass_g) %>% 
  miss_var_summary()
```
These results make sense biologically since mammals have 100% missing for egg mass which is correct as mammals do not lay eggs. NAs in this data set may not all just be missing data but also serve as placeholders for groups were that data variable does not apply as well. 


**9. The `amphibio` data have variables that classify species as fossorial (burrowing), terrestrial, aquatic, or arboreal.Calculate the number of NA's in each of these variables. Do you think that the authors intend us to think that there are NA's in these columns or could they represent something else? Explain.**
```{r}
amphibo %>% 
  select(Fos, Ter, Aqu, Arb) %>% 
  miss_var_summary()
```
I believe that author did not intend to use the NAs in these varibales as missing values but rather as 0 values. When looking at the pdf it states that these variables are binary and these missing values should in reality mean 0. Now it appears that data is missing but I think the author was trying to say that the NAs represent species that do not fall into a particular category. 



**10. Now that we know how NA's are represented in the `amniota` data, how would you load the data such that the values which represent NA's are automatically converted?**
```{r}
amniota_new <- read_csv(file = "data/amniota.csv", na = c("NA", " ", ".", "-999")) %>% 
  clean_names()
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  