---
title: "dplyr Superhero"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    toc: true
    toc_float: true
    keep_md: true
---

## Learning Goals  
*At the end of this exercise, you will be able to:*    
1. Develop your dplyr superpowers so you can easily and confidently manipulate dataframes.  
2. Learn helpful new functions that are part of the `janitor` package.  

## Instructions
For the second part of lab today, we are going to spend time practicing the dplyr functions we have learned and add a few new ones. This lab doubles as your homework. Please complete the lab and push your final code to GitHub.  

## Load the libraries
```{r}
library("tidyverse")
library("janitor")
```

## Load the superhero data
These are data taken from comic books and assembled by fans. The include a good mix of categorical and continuous data.  Data taken from: https://www.kaggle.com/claudiodavi/superhero-set  

Check out the way I am loading these data. If I know there are NAs, I can take care of them at the beginning. But, we should do this very cautiously. At times it is better to keep the original columns and data intact.  
```{r}
superhero_info <- read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))
```

## Data tidy
1. Some of the names used in the `superhero_info` data are problematic so you should rename them here. Before you do anything, first have a look at the names of the variables. You can use `rename()` or `clean_names()`.    

## `tabyl`
The `janitor` package has many awesome functions that we will explore. Here is its version of `table` which not only produces counts but also percentages. Very handy! Let's use it to explore the proportion of good guys and bad guys in the `superhero_info` data.  
```{r}
tabyl(superhero_info, Alignment)
```

1. Who are the publishers of the superheros? Show the proportion of superheros from each publisher. Which publisher has the highest number of superheros?  

```{r}
superhero_info %>% 
  tabyl(Publisher)
```

Marvel comics has the highest number of publisher. 


2. Notice that we have some neutral superheros! Who are they? List their names below.  

```{r}
neutral_name <- superhero_info %>% 
  filter(Alignment == "neutral")
neutral_name$name
```


## `superhero_info`
3. Let's say we are only interested in the variables name, alignment, and "race". How would you isolate these variables from `superhero_info`?

```{r}
superhero_info %>% 
  select(name,Alignment,Race)
```


## Not Human
4. List all of the superheros that are not human.

```{r}
not_human <- superhero_info %>% 
  filter(Race != "Human")
  not_human$name
```

## Good and Evil
5. Let's make two different data frames, one focused on the "good guys" and another focused on the "bad guys".

```{r}
good_guys <- superhero_info %>% 
    filter(Alignment == "good")
```

```{r}
bad_guys <- superhero_info %>% 
    filter(Alignment == "bad")
```


6. For the good guys, use the `tabyl` function to summarize their "race".

```{r}
good_guys %>% 
  tabyl(Race)
```


7. Among the good guys, Who are the Vampires?

```{r}
vampire_good_guys <- good_guys %>% 
  filter(Race == "Vampire")
vampire_good_guys$name
```


8. Among the bad guys, who are the male humans over 200 inches in height?

```{r}
tall_bad_guys <- bad_guys %>% 
  filter(Height >200)
tall_bad_guys$name
```

9. Are there more good guys or bad guys with green hair? 
```{r}
table(good_guys$`Hair color`)
```

```{r}
table(bad_guys$`Hair color`)
```
There are 7 good guys with green hair and only 1 bad guy with green hair meaning there are more green colored hair in the good guys group. 


10. Let's explore who the really small superheros are. In the `superhero_info` data, which have a weight less than 50? Be sure to sort your results by weight lowest to highest.

```{r}
superhero_info %>% 
  filter(Weight <50) %>% 
  arrange((Weight))
```


11. Let's make a new variable that is the ratio of height to weight. Call this variable `height_weight_ratio`.    

```{r}
update_superhero_table <- superhero_info %>% 
  mutate(height_weight_ratio = Height/Weight) 
  
```


12. Who has the highest height to weight ratio?  

```{r}
update_superhero_table %>% 
  arrange(desc(height_weight_ratio))
```

Groot has the largest height to weight ratio with a height of 701.0 and weight of 4 giving a ratio of 175.25

## `superhero_powers`
Have a quick look at the `superhero_powers` data frame.  

```{r}
glimpse(superhero_powers)
```


13. How many superheros have a combination of agility, stealth, super_strength, stamina?

```{r}
filtered_powers <- superhero_powers %>% 
  select(hero_names,Agility, Stealth, `Super Strength`, Stamina) %>% 
  filter(Agility == TRUE & Stealth == TRUE & `Super Strength` ==  TRUE & Stamina == TRUE)
filtered_powers$hero_names

```
There are a total of 40 superheros with all those powers. 

## Your Favorite
14. Pick your favorite superhero and let's see their powers!  

```{r}
superhero_powers %>% 
  filter(hero_names == "Iron Man") %>% 
  select_if(all)
```


15. Can you find your hero in the superhero_info data? Show their info! 

```{r}
superhero_info %>% 
  filter(name == "Iron Man")
```


## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  